A Comparision of Bootstrap Confidence Interval Methods

### 2022-12-06 project/PaperTemplate/paper.Rmd
Last commit before deleting old simulation
Knitted to pdf

```{r find CIs}
findCIs <- function(sampleinput, nboot) {

  sample_mean <- mean(~ X, data = sampleinput)
  sample_sd <- sd(~ X, data = sampleinput)
  sample_n <- nrow(sampleinput)
  sample_se <- sample_sd/sqrt(sample_n)

  alphalow <- 0.025
  alphahigh <- 0.975

  #perform bootstrap
  mean <- rep(0,nboot)
  sd <- rep(0,nboot)
  se <- rep(0,nboot)
  t_stat <- rep(0,nboot)

  for (i in 1:nboot) {
    resampled <- as.data.frame(mosaic::resample(sampleinput, replace = TRUE))
    mean[i] <- mean(~ X, data = resampled)
    sd[i] <- sd(~ X, data = resampled)
    se[i] <- sd[i]/sqrt(nrow(resampled))
    t_stat[i] <- (mean[i] - sample_mean)/se[i]
  }

  boot_means <- as.data.frame(mean)
  boot_t_stat <- as.data.frame(t_stat)

  # find the standard confidence interval
  standard_interval <- sample_mean + 
    qnorm(c(0.025, 0.975)) * sample_sd/sqrt(sample_n)

  # find percentile interval
  percentile_interval <- qdata(~ mean, c(0.025, 0.975), data = boot_means)

  # find BC interval
  less_than_sample_mean <- sum(ifelse(boot_means <= sample_mean, 1, 0))/nboot
  z0 <- qnorm(less_than_sample_mean)
  newlow <- pnorm(2*z0 + qnorm(alphalow))
  newhigh <- pnorm(2*z0 + qnorm(alphahigh))

  bc_interval <- qdata(~ mean, c(newlow, newhigh), data = boot_means)

  # find BCa interval
  theta <- function(x){mean(x)}
  jackknife_results <- bootstrap::jackknife(sampleinput$X, theta)
  jackmean <- mean(~ jackknife_results$jack.values)
  estimated_a <- (1/6)*sum((jackknife_results$jack.values - jackmean)^3)/
    (sum((jackknife_results$jack.values - jackmean)^2))^(1.5)
  a_newlow <- pnorm(z0 + (z0 + qnorm(alphalow))/
                      (1-estimated_a*(z0 + qnorm(alphalow))))
  a_newhigh <- pnorm(z0 + (z0 + qnorm(alphahigh))/
                       (1-estimated_a*(z0 + qnorm(alphahigh))))
  bca_interval <- qdata(~ mean, c(a_newlow, a_newhigh), data = boot_means)

  # find studentized interval
  t_q <- unname(quantile(boot_t_stat$t_stat, c(0.025, 0.975)))
  lower_q <- t_q[1]
  upper_q <- t_q[2]
  t_interval <- c(sample_mean - (upper_q*sample_se), sample_mean - (lower_q*sample_se))

  output <- list(standard_interval, percentile_interval, bc_interval, 
                 bca_interval, t_interval)
  names(output) <- c("Standard Interval", "Percentile Interval", "BC Interval", 
                     "BCa Interval", "Studentized Interval")
  output
}
```

@@

```{r run simulation normal}
run_simulation_normal <- function(simulation_rep){
  trueparameter <- 0

  X <- rnorm(sample_sizes, mean = 0, sd = 1)
  values <- data.frame(X)
  output <- findCIs(values, 2000)
  ci <- matrix(c(trueparameter >= output$`Standard Interval`[1] & 
    trueparameter <= output$`Standard Interval`[2],
    trueparameter >= output$`Percentile Interval`[1] & 
    trueparameter <= output$`Percentile Interval`[2],
    trueparameter >= output$`BC Interval`[1] &
    trueparameter <= output$`BC Interval`[2],
    trueparameter >= output$`BCa Interval`[1] & 
    trueparameter <= output$`BCa Interval`[2],
    trueparameter >= output$`Studentized Interval`[1] & 
    trueparameter <= output$`Studentized Interval`[2]), nrow = 1)
  res <- rbind(data.frame(), ci)
  res
}

@@

```{r run simulation gamma}
run_simulation_gamma <- function(simulation_rep){
  trueparameter <- 4

  X <- rgamma(sample_sizes, shape = 1, scale = 4)
  values <- data.frame(X)
  output <- findCIs(values, 2000)
  ci <- matrix(c(trueparameter >= output$`Standard Interval`[1] & 
    trueparameter <= output$`Standard Interval`[2],
    trueparameter >= output$`Percentile Interval`[1] & 
    trueparameter <= output$`Percentile Interval`[2],
    trueparameter >= output$`BC Interval`[1] &
    trueparameter <= output$`BC Interval`[2],
    trueparameter >= output$`BCa Interval`[1] & 
    trueparameter <= output$`BCa Interval`[2],
    trueparameter >= output$`Studentized Interval`[1] & 
    trueparameter <= output$`Studentized Interval`[2]), nrow = 1)
  res <- rbind(data.frame(), ci)
  res
}

```{r run simulation log normal}

#Third, we sample from a Log-Normal distribution with mean log of 0 and a mean standard deviation of 1. Therefore, the true population mean of this distribution is $e^{\frac{1}{2}}$ [@Guillaume21].

run_simulation_log_normal <- function(simulation_rep){
  trueparameter <- exp(1/2)

  X <- rlnorm(sample_sizes, meanlog = 0, sdlog = 1)
  values <- data.frame(X)
  output <- findCIs(values, 2000)
  ci <- matrix(c(trueparameter >= output$`Standard Interval`[1] & 
    trueparameter <= output$`Standard Interval`[2],
    trueparameter >= output$`Percentile Interval`[1] & 
    trueparameter <= output$`Percentile Interval`[2],
    trueparameter >= output$`BC Interval`[1] &
    trueparameter <= output$`BC Interval`[2],
    trueparameter >= output$`BCa Interval`[1] & 
    trueparameter <= output$`BCa Interval`[2],
    trueparameter >= output$`Studentized Interval`[1] & 
    trueparameter <= output$`Studentized Interval`[2]), nrow = 1)
  res <- rbind(data.frame(), ci)
  res
}

```{r run simulation beta}
run_simulation_beta <- function(simulation_rep){
  trueparameter <- 0.8

  X <- rbeta(sample_sizes, shape1 = 4, shape2 = 1)
  values <- data.frame(X)
  output <- findCIs(values, 2000)
  ci <- matrix(c(trueparameter >= output$`Standard Interval`[1] & 
    trueparameter <= output$`Standard Interval`[2],
    trueparameter >= output$`Percentile Interval`[1] & 
    trueparameter <= output$`Percentile Interval`[2],
    trueparameter >= output$`BC Interval`[1] &
    trueparameter <= output$`BC Interval`[2],
    trueparameter >= output$`BCa Interval`[1] & 
    trueparameter <= output$`BCa Interval`[2],
    trueparameter >= output$`Studentized Interval`[1] & 
    trueparameter <= output$`Studentized Interval`[2]), nrow = 1)
  res <- rbind(data.frame(), ci)
  res
}

@@

```{r collect data gamma}
collect_data_gamma <- function(sim_reps) {
  data2 <- matrix(nrow = 7, ncol = 6)
  sample_sizes = 4

  for (i in 1:7) {
    simulation_rep <- seq(1,sim_reps)
    results <- mclapply(simulation_rep, run_simulation_gamma, 
                        mc.cores = detectCores())
    df <- data.frame(matrix(unlist(results), nrow=length(results), byrow=TRUE))
    ci_prop <- colSums(df)/sim_reps

    data2[i, 1] <- sample_sizes
    data2[i, 2:6] <- ci_prop
    sample_sizes <- sample_sizes*2
  }

  gamma_output <- as.data.frame(data2) %>% 
    rename("Sample_Size" = V1, "Standard" = V2, "Percentile" = V3, "BC" = V4, 
           "BCa" = V5, "Studentized" = V6) 
  gamma_output
}

```{r collect data log normal}
collect_data_log_normal <- function(sim_reps) {
  data2 <- matrix(nrow = 7, ncol = 6)
  sample_sizes = 4

  for (i in 1:7) {
    simulation_rep <- seq(1,sim_reps)
    results <- mclapply(simulation_rep, run_simulation_log_normal, 
                        mc.cores = detectCores())
    df <- data.frame(matrix(unlist(results), nrow=length(results), byrow=TRUE))
    ci_prop <- unname(colSums(df)/sim_reps)

    data2[i, 1] <- sample_sizes
    data2[i, 2:6] <- ci_prop
    sample_sizes <- sample_sizes*2
  }

  log_normal_output <- as.data.frame(data2) %>% 
    rename("Sample_Size" = V1, "Standard" = V2, "Percentile" = V3, "BC" = V4, 
           "BCa" = V5, "Studentized" = V6) 
  log_normal_output
}
```

```{r collect data normal}
collect_data_normal <- function(sim_reps) {
  data2 <- matrix(nrow = 7, ncol = 6)
  sample_sizes = 4

  for (i in 1:7) {
    simulation_rep <- seq(1,sim_reps)
    results <- mclapply(simulation_rep, run_simulation_normal, 
                        mc.cores = detectCores())
    df <- data.frame(matrix(unlist(results), nrow=length(results), byrow=TRUE))
    ci_prop <- unname(colSums(df)/sim_reps)

    data2[i, 1] <- sample_sizes
    data2[i, 2:6] <- ci_prop
    sample_sizes <- sample_sizes*2
  }

  normal_output <- as.data.frame(data2) %>% 
    rename("Sample_Size" = V1, "Standard" = V2, "Percentile" = V3, "BC" = V4, 
           "BCa" = V5, "Studentized" = V6) 
  normal_output
}

```{r collect data beta}
collect_data_beta <- function(sim_reps) {
  data2 <- matrix(nrow = 7, ncol = 6)
  sample_sizes = 4

  for (i in 1:7) {
    simulation_rep <- seq(1,sim_reps)
    results <- mclapply(simulation_rep, run_simulation_beta, 
                        mc.cores = detectCores())
    df <- data.frame(matrix(unlist(results), nrow=length(results), byrow=TRUE))
    ci_prop <- unname(colSums(df)/sim_reps)

    data2[i, 1] <- sample_sizes
    data2[i, 2:6] <- ci_prop
    sample_sizes <- sample_sizes*2
  }

  beta_output <- as.data.frame(data2) %>% 
    rename("Sample_Size" = V1, "Standard" = V2, "Percentile" = V3, "BC" = V4, 
           "BCa" = V5, "Studentized" = V6) 
  beta_output
}

@@

```{r, cache = TRUE, eval = FALSE}
set.seed(495)
sim_reps <- 10000

gamma_data_4_128 <- collect_data_gamma(sim_reps)
write.csv(gamma_data_4_128,"gamma_data_4_128.csv", row.names = FALSE)

normal_data_4_128 <- collect_data_normal(sim_reps)
write.csv(normal_data_4_128,"normal_data_4_128.csv", row.names = FALSE)

log_normal_data_4_128 <- collect_data_log_normal(sim_reps)
write.csv(log_normal_data_4_128,"log_normal_data_4_128.csv", row.names = FALSE)

beta_data_4_128 <- collect_data_beta(sim_reps)
write.csv(beta_data_4_128,"beta_data_4_128.csv", row.names = FALSE)
```

```{r, include = FALSE, eval = FALSE}

lognormal_df_long <- lognormal_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

lognormal_fig <- ggplot(data = lognormal_df_long, 
                    aes(x = Sample_Size, y = prop, color = method)) + 
  geom_point() + geom_line() + 
  labs(title = "Proportion of CIs that Contain the True Parameter from Log-Normal 
       Distibution", x = "Sample Size", 
       y = "Proportion (of 10,000 Simulations)", color = "Method") +
  theme_light() + 
  theme(legend.position="bottom")

lognormal_table <- lognormal_df %>%
  kable(booktabs = TRUE, align = "c", caption = "log normal", 
        col.names = gsub("[_]", " ", names(lognormal_df)))

lognormal_df <- read.csv(file = 'log_normal1_data_10000.csv')

@@

gamma_df <- read.csv(file = 'gamma_data1_10000.csv')
normal_df <- read.csv(file = 'normal_data1_10000.csv')
log_normal_df <- read.csv(file = 'log_normal1_data_10000.csv')

gamma_df_long <- log_normal_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

normal_table <- normal_df %>%
  kable(booktabs = TRUE, align = "c", caption = "Sampled from Normal Dist.", 
        col.names = gsub("[_]", " ", names(normal_df)))
```

```{r}
gamma_fig
gamma_table
```

@@

```{r}
normal_fig
normal_table
```

@@

```{r}
master_df_long <- master_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

master_fig
```
```{r, warning = FALSE}
library(boot)

@@

```{r}
set.seed(495)
log_normal_df <- run_sim_log_normal(10000)
write.csv(log_normal_df,"log_normal_df.csv", row.names = FALSE)

@@

>>>>

@@

@@

@@

@@

@@

@@

@@

@@

```{r, warning = FALSE, eval = FALSE}
set.seed(495)
log_normal_df <- run_sim_log_normal(10000)
write.csv(log_normal_df,"log_normal_df.csv", row.names = FALSE)

@@

gamma_df <- read.csv(file = 'gamma_data_df.csv')
normal_df <- read.csv(file = 'normal_data_df.csv')
log_normal_df <- read.csv(file = 'log_normal_df.csv')

gamma_df_long <- gamma_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

normal_table <- normal_df %>%
  kable(booktabs = TRUE, align = "c", caption = "Sampled from Normal Dist.", 
        col.names = gsub("[_]", " ", names(normal_df)))

log_normal_df_long <- log_normal_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

log_normal_fig <- ggplot(data = log_normal_df_long, 
                    aes(x = Sample_Size, y = prop, color = method)) + 
  geom_point() + geom_line() + 
  labs(title = "Proportion of CIs that Contain the True Parameter from Log-Normal 
       Distibution", x = "Sample Size", 
       y = "Proportion (of 10,000 Simulations)", color = "Method",
       caption = "Fig. 8: Graph showing the Accuracy of Bootstrap CI Methods for different sample sizes on Log-Normal Dist.") +
  theme_light() +
  theme(legend.position="bottom")

log_normal_table <- log_normal_df %>%
  kable(booktabs = TRUE, align = "c", caption = "Sampled from Normal Dist.", 
        col.names = gsub("[_]", " ", names(log_normal_df)))

```{r}
log_normal_fig
log_normal_table

normal_fig
normal_table

gamma_fig
gamma_table

```{r, eval = FALSE}

normal_df <- read.csv(file = 'normal_data_df.csv')

master_df_long <- master_df %>%
  pivot_longer(cols = 2:6, names_to = "method", values_to = "prop")

master_fig
```
